\paperTitle

### 2014-01-31 doc/icml2014/directed.tex
push

@@

\begin{figure}
  \centering
  \subfigure[Hidden Markov Model] {
    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
    \label{fig:examples-hmm}
  }
%  \subfigure[Directed grid model] {
%    \label{fig:examples-grid}
%    \includegraphics{figures/grid.pdf}
%  }
  \subfigure[Tree model] {
    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
    \label{fig:examples-tree}
  }
  \caption{Additional examples of models learnable using \LearnMarginals}
  \label{fig:examples}
\end{figure}

\paragraph{Hidden Markov Model}

In this example (\figureref{examples-hmm}), we assume that
  $\Pr(x_i|h_i) = O  ~\forall i$  and that $\Pr(h_{i+1} | h_i)
  = T ~\forall i$ (i.e. we have parameter sharing).
Note that while the first (and last) hidden variables $h_1, h_T$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_T$ respectively) whose parameters we know because they share
  parameters, $O$.
In step 1 of our algorithm, we use the bottleneck $h_2$ with views $x_1,
  x_2, x_3$ and solve $O$.
In step 2 of our algorithm, we can recover $\pi$ by solving for the
  unary clique $\{h_1\}$ and recover $T$ from the clique $\{h_{1},
  h_{2}\}$.

\paragraph{Latent Tree Structure}

In the latent tree structure (\figureref{examples-tree}), let the
  parameters be $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T ~i \in \{2,3,4\}$
  and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O ~i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.

In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
  one of views (they are all exclusive). 
To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
  $h_1$.

>>>>

\paragraph{Example: hidden Markov model.}

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$ (i.e. we have parameter sharing).
While the first (and last) hidden variables $h_1, h_T$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_T$ respectively) whose parameters we know because they share
  parameters, $O$.
%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.

\paragraph{Example: latent tree model.}

In the latent tree model (\figureref{examples-tree}), the parameters
are $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T$ for $i \in \{2,3,4\}$
and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O$ for $i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$).
  %Note that while
  %$x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
  %$h_1$.

%In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
%  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
%  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
%In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
%  one of views (they are all exclusive). 
%To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
%  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
%  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
%  $h_1$.

@@

\begin{figure}
  \centering
  \subfigure[Hidden Markov Model] {
    \includegraphics[width=0.45\columnwidth]{figures/hmm.pdf}
    \label{fig:examples-hmm}
  }
%  \subfigure[Directed grid model] {
%    \label{fig:examples-grid}
%    \includegraphics{figures/grid.pdf}
%  }
  \subfigure[Tree model] {
    \includegraphics[width=0.45\columnwidth]{figures/tree.pdf}
    \label{fig:examples-tree}
  }
  \caption{Examples of directed models satisfying the exclusive views property.}
  \label{fig:examples}
\end{figure}

### 2014-01-31 doc/icml2014/discussion.tex
push

Parameter estimation for latent variable models using the method of
  moments has received much attention in last few years.
Our work uses the \TensorFactorize algorithm presented by
  \citet{anandkumar13tensor}, though \citet{anandkumar12moments} first
  studied the consistent parameter estimation algorithm for the
  three-view mixture model.
%\todo{What about Mossel and Roch?}
Several authors have extended this work to settings that do not meet the
  criterion \propertyref{bottleneck}, but rely on knowing more about
  the underlying correlation structure of the problem.
\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar2013linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered. 
In comparison, our work applies to a general family of models, including
  latent tree models, directed and undirected grids, with no assumptions
  on the correlation structure aside from independences encoded in the
  graph.

@@

In this paper, we've described an algorithm that can consistently
  estimate parameters for the class of directed and undirected graphical
  models for which each hidden variable has at least three conditionally
  independent observed variables (\propertyref{bottleneck}).
This condition is sufficiently general to capture a variety of popular
  models like latent trees and grids (like a pairwise latent Markov
  random field).
The undirected model setting applies to log-linear models, allowing
  for discriminative featurization; furthermore, in practice, the
  bottleneck condition can be enforced on a model if the observed
  features can be split into two uncorrelated sets.

The key property we exploited was the presence of exclusive views
  (\propertyref{exclusive-views}) with learnable parameters.
In some sense, the presence of these views helps tie down the hidden
  variables to the data. 
While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
  halpern13noisyor} are able to learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  \propertyref{bottleneck}.
Thus, while \propertyref{bottleneck} is much stronger than
  \propertyref{exclusive-views}, we expect that without any knowledge of
  the correlation between variables, \propertyref{bottleneck} is
  necessary for identifiability.

>>>>

% Summarize paper
For latent-variable models,
there exists a tension between
local optimization of likelihood,
which is broadly applicable but has no theoretical guarantees,
and the method of moments, which provides consistent estimators
but are limited to models with special structure.
The purpose of this work is to show that the two methods
can be used synergistically to produce consistent estimates
for a broader class of models.
The key idea is to (i) perform tensor factorization \citep{anandkumar13tensor} to
obtain information about each hidden variable in the form of conditional moments;
(ii) use composite likelihood to estimate the marginal distributions of latent variables;
and (iii) in the case of undirected models,
use the marginal distributions to formulate a convex optimization to
recover the model parameters.

% Bottleneck is key
Our approach allows us to handle a broad class of directed and undirected
models in which each hidden variable is a ``bottleneck''---that is,
it has three conditionally independent observations,
(although the dependence on the observations could be quite complex as a function of the parameters).
This bottleneck property of \citep{anandkumar13tensor}
has been exploited in many other contexts,
including Latent Dirichlet Allocation \cite{anandkumar12lda},
parsing \cite{hsu12identifiability},
noisy-or Bayesian networks \cite{halpern13noisyor},
mixture of linear regressions \cite{chaganty13regression},
and others.

%Much recent work on method of moments for estimating latent-variable models
%has focused on increasing applicability to broader range of models,

%In this paper, we we've described an algorithm that can consistently
%  estimate parameters for the class of directed and undirected graphical
%  models for which each hidden variable has at least three conditionally
%  independent observed variables (\propertyref{bottleneck}).

%This condition is sufficiently general to capture a variety of popular
  %models like latent trees and grids (like a pairwise latent Markov random field).
%The undirected model setting applies to log-linear models, allowing
%  for discriminative featurization; furthermore, in practice, the
%  bottleneck condition can be enforced on a model if the observed
%  features can be split into two uncorrelated sets.

% Classes of models
Notably, the bottleneck property allows us to perform consistent parameter
estimation in undirected log-linear models with hidden variables with high
treewidth, which

The key property we exploited was the presence of exclusive views
  (\propertyref{exclusive-views}) with learnable parameters.
In some sense, the presence of these views helps tie down the hidden
  variables to the data. 
While other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
  halpern13noisyor} are able to learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  \propertyref{bottleneck}.
% TOO BOLD
%Thus, while \propertyref{bottleneck} is much stronger than
%  \propertyref{exclusive-views}, we expect that without any knowledge of
%  the correlation between variables, \propertyref{bottleneck} is
%  necessary for identifiability.

% Broad interest in learning more models
The use of method of moments for learning latent-variable models
%\cite{balle11transducer}
including 
Much of the Parameter estimation for latent-variable models using the method of
  moments has received much attention in recent years,
aimed at increasing their applicability to broader classes of models.

Our work uses the \TensorFactorize algorithm presented by
  \citet{anandkumar13tensor}, first
  studied the consistent parameter estimation algorithm for the
  three-view mixture model.

%\todo{What about Mossel and Roch?}

\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar2013linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered.

work to settings that do not meet the
  criterion \propertyref{bottleneck}, but rely on knowing more about
  the underlying correlation structure of the problem.

In comparison, our work applies to a general family of models, including
  latent tree models, directed and undirected grids, with no assumptions
  on the correlation structure aside from independences encoded in the
  graph.

@@

% Identifiability
Of course, there are models which 
\cite{kruskal77three}

### 2014-01-31 doc/icml2014/examples.tex
Removed false intuitions about the compositie-likelihood
push

%\subsection{More examples}

@@

\paragraph{Aggregating observations}
We describe two practical considerations to use samples more efficiently.
Firstly, if we have multiple exclusive views for a hidden variable,
  intuitively, it is better to aggregate over them. 
For example, consider a hidden variable $h_1$ with multiple exclusive
  views $x_1$ and $x_2$.
With the method of moments perspective, to learn the marginal
  distribution over $h_1$, one must solve the following reconstruction
  problem, 
\begin{align*}
  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \half \|Z_{h_1}(\mOpp{1}{1}) - M_1 \|^2 + \half \|Z_{h_1}(\mOpp{2}{1}) - M_2 \|^2,
\end{align*}
which does not have a closed form solution. 
In contrast, this naturally fits into the convex optimization framework, where $\hat Z_{h_1}$ will now be,
\begin{align*}
  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \sum_{\vx \in \sD} \log Z_{h_1}( \mOpp{1}{1}[x_1] \odot \mOpp{1}{1}[x_1] ),
\end{align*}
where $\cdot$ denotes element-wise multiplication.
An important note to make is that \TensorFactorize only returns
  a solution up to permutation; if $\mOpp{1}{1}$ and
  $\mOpp{1}{2}$ do not belong to the same bottleneck (e.g. $x^a_2$ and
  $x^b_2$ in \figureref{examples-tree}), then some
  care must be taken to ensure they have the same labelling.

Secondly, we can exploit parameter sharing by aggregating over
  disjoint sets of observed variables. 
For example, in \figureref{examples-hmm}, we can aggregate the statistics for
  bottlenecks $h_i$ with views $\{x_{i-1}, x_{i}, x_{i+1}\}$ before
  running \TensorFactorize; this will give us a consistent estimate for
  $O$ (as well as $T$).

>>>>

@@

%\paragraph{Aggregating observations}
%We describe two practical considerations to use samples more efficiently.
%Firstly, if we have multiple exclusive views for a hidden variable,
%  intuitively, it is better to aggregate over them. 
%For example, consider a hidden variable $h_1$ with multiple exclusive
%  views $x_1$ and $x_2$.
%With the method of moments perspective, to learn the marginal
%  distribution over $h_1$, one must solve the following reconstruction
%  problem, 
%\begin{align*}
%  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \half \|Z_{h_1}(\mOpp{1}{1}) - M_1 \|^2 + \half \|Z_{h_1}(\mOpp{2}{1}) - M_2 \|^2,
%\end{align*}
%which does not have a closed form solution. 
%In contrast, this naturally fits into the convex optimization framework, where $\hat Z_{h_1}$ will now be,
%\begin{align*}
%  \hat Z_{h_1} &= \arg\min_{Z_{h_1}} \sum_{\vx \in \sD} \log Z_{h_1}( \mOpp{1}{1}[x_1] \odot \mOpp{1}{1}[x_1] ),
%\end{align*}
%where $\cdot$ denotes element-wise multiplication.
%An important note to make is that \TensorFactorize only returns
%  a solution up to permutation; if $\mOpp{1}{1}$ and
%  $\mOpp{1}{2}$ do not belong to the same bottleneck (e.g. $x^a_2$ and
%  $x^b_2$ in \figureref{examples-tree}), then some
%  care must be taken to ensure they have the same labelling.
%
%Secondly, we can exploit parameter sharing by aggregating over
%  disjoint sets of observed variables. 
%For example, in \figureref{examples-hmm}, we can aggregate the statistics for
%  bottlenecks $h_i$ with views $\{x_{i-1}, x_{i}, x_{i+1}\}$ before
%  running \TensorFactorize; this will give us a consistent estimate for
%  $O$ (as well as $T$).

### 2014-01-31 doc/icml2014/undirected.tex
push

\paragraph{Pseudolikelihood.}
Unlike for directed models, in the undirected case,
graphs with high tree-width present a challenge
for even fully-supervised maximum likelihood.
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo},
which optimizes the likelihood of each variable $a$ given its neighbors $\sN(a)$:
$\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})$.
In the fully-supervised setting, pseudolikelihood provides consistent estimates,
albeit at the cost of statistical efficiency.
However, if we write the objective, we will see that the log-partition function $A(\theta)$
is replaced with a sum over $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
which cannot be computed from $Z_\sC$, since the neighbors $\sN(a)$ in general do not form a clique.
However, we adapt our earlier algorithm to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
Of course, this method is only viable for graphs of low degree ($|\sN(a)|$ is small).
For example, in the grid model, the degree is only 4.
Note that neither sample nor computational complexity depends on the treewidth.

>>>>

\paragraph{Pseudolikelihood.}
Unlike for directed models, in the undirected case,
graphs with high tree-width present a challenge
for even fully-supervised maximum likelihood.
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo},
which optimizes the likelihood of each variable $a$ given its neighbors $\sN(a)$:
$\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})$.
In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
albeit at the cost of statistical efficiency
(coincidentally, the same high-level motivation for using method of moments in the first place).
However, if we write the objective, we will see that the log-partition function $A(\theta)$
is replaced with a sum over $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
which cannot be computed from $Z_\sC$, since the neighbors $\sN(a)$ in general do not form a clique.
However, we adapt our earlier algorithm to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
Of course, this method is only viable for graphs of low degree ($|\sN(a)|$ is small).
For example, in the grid model, the degree is only 4.
Note that neither sample nor computational complexity depends on the treewidth.
